---
title: "Understanding Toronto's Red Light Camera Data: Formats, Accessibility, and Implications for Traffic Safety"
author: "Duanyi Su"
date: "September 23, 2024"
format: pdf
---


## Introduction
Red light cameras play a crucial role in reducing traffic violations and enhancing road safety in urban areas. In Toronto, these cameras are installed at key intersections to monitor traffic signal compliance, serving as an effective tool for improving road safety. Analyzing the data collected from these cameras can provide valuable insights into traffic patterns, enforcement efficiency, and overall public safety. The availability of different data formats and regular updates is a testament to the city's commitment to transparency and data-driven decision-making. This paper examines the metadata of the red light camera datasets provided by Open Data Toronto, focusing on data formats, update frequencies, and their implications for data accessibility and usability.
The remainder of this paper is organized as follows: the data section discusses the dataset's source, structure, and ethical considerations, followed by an analysis using visualizations and tables. We conclude with recommendations and suggestions for future research.


## Data

### Data Source and Context
The dataset analyzed was sourced from Open Data Toronto's "Red Light Cameras Data" collection. It includes metadata on different file formats (GeoJSON, CSV, GPKG) and their respective update frequencies. This metadata provides insights into how the city shares traffic safety data, making it accessible to various stakeholders such as researchers, policymakers, and the general public.
From an ethical perspective, sharing red light camera data raises concerns about privacy and surveillance. However, the benefits of improving traffic safety and compliance outweigh these concerns. The diverse data formats suggest a strong effort toward accessibility and transparency, enabling a wide range of data users to engage with the information.

### Data Summary and Analysis
The following table summarizes the metadata entries for Toronto's red light camera data:
```{r}
library(knitr)
library(dplyr)

# Loading the data into R and displaying it using knitr::kable
red_light_data <- read.csv("Red Light Cameras Data - 4326.csv")
red_light_data_summary <- red_light_data %>% select(name, format, last_modified)
kable(red_light_data_summary, caption = "Table 1: Summary of Red Light Camera Data Metadata")
```
As shown in Table 1, the dataset offers multiple formats and regular updates, indicating a commitment to providing up-to-date information for traffic safety analysis.

### Visualization
The graph below illustrates the distribution of data formats available in the dataset, emphasizing the variety of formats offered:
```{r}
library(ggplot2)

# Creating a bar plot using ggplot2
ggplot(red_light_data, aes(x = format)) +
  geom_bar(fill = "grey") +
  theme_minimal() +
  labs(title = "Figure 1: Distribution of Data Formats for Red Light Camera Data",
       x = "Data Format", y = "Count")
```
Figure 1 demonstrates that multiple data formats are provided, ensuring that different user groups can access and utilize the data effectively.


## References
- Wickham, H. (2016). *ggplot2: Elegant Graphics for Data Analysis*. Springer-Verlag New York.
- Xie, Y. (2023). *knitr: A General-Purpose Package for Dynamic Report Generation in R*. R package version.
- Iannone, R., Cheng, J., & Schloerke, B. (2022). *gt: Easily Create Presentation-Ready Display Tables*. R package version.
- Open Data Toronto. (n.d.). Red Light Cameras Data. Retrieved from https://open.toronto.ca/dataset/red-light-cameras/


## Acknowledgments
We acknowledge Open Data Toronto for providing the red light camera data. For more details, visit our GitHub repository: [GitHub Link](https://github.com/DuanyiSu/sta304-paper1)
